\documentclass{essay}

% Document metadata
\title{Sample Essay Using the Essay Template}
\author{Jane Doe}
\date{\today}

\begin{document}

% Title page with no header/footer
\maketitle

% Abstract or frontmatter on first page
\begin{abstract}
This is a sample abstract demonstrating the minimal LaTeX essay template. The template provides a clean, professional look with dark blue colors for section headings and hyperlinks. It includes proper formatting for academic essays with headers, footers, and bibliography support.
\end{abstract}

\section{Introduction}

This document demonstrates the essay template class. The template is designed to be minimal yet complete, providing all the essential features needed for academic essays.
It is published under GPL-3.0, its source is available on GitHub.\footnote{\href{https://github.com/jankapunkt/latex-essay}{https://github.com/jankapunkt/latex-essay}}

The template features:
\begin{itemize}
    \item A4 paper format with two-sided layout
    \item 1cm margins on left and right sides
    \item 11pt font size for readability
    \item 1.15 line height for comfortable reading
    \item Dark blue colors for headlines and hyperlinks
    \item Headers showing section titles
    \item Footers with centered page numbers
    \item Clean title page without header or footer
\end{itemize}

\section{Main Content}

This section demonstrates the header functionality. Notice how the section title appears in the header at the top of the page. The page number is centered in the footer.

\subsection{Subsection Example}

Subsections are also formatted in dark blue, maintaining visual consistency throughout the document. The hierarchical structure is clear and easy to follow.

\subsubsection{Subsubsection Example}

Even third-level headings maintain the dark blue color scheme, ensuring a cohesive look throughout the document.

\section{Citations and References}

The template supports BibTeX citations using the IEEE transaction style (ieeetr). You can cite references like this \cite{example2025} and they will appear in the bibliography at the end of the document.

Multiple citations are also supported \cite{example2024, another2023}.

\section{Conclusion}

This template provides a solid foundation for academic essays, combining clean design with practical functionality. The dark blue color scheme adds visual interest while maintaining professionalism.

% Bibliography
\bibliography{references}

\end{document}
